### 2024-04-02 paper/paper.qmd
initial set up

---
title: "Data evidence of financial globalization: how excess saving in East Asian affect US interest rate"
author: Yingzhi Zhang
thanks: "Code and data are available at: https://github.com/melissazhangyz/east_asia_finance."
date: today
date-format: long
abstract: "TO BE UPDATED"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction
The global economy has been undergoing a lower interest rate for a while. Though the lack of confidence in investment shown by a global saving glut is a strong explanation, the increasing saving rate is not observed in developed countries but only developing countries. Such differentiation in economies around the world leads to an analysis of differentiation of countries and their roles in global capital markets. This study uses linear regression and ANOVA analysis to examine the correlation between US interest rate and saving rate in middle-and low-income countries in East Asia and Pacific. A negative correlation shows strong power of developing countries, which ultimately explains the differences between previous and current structure differences of global capital market.

We use @citeR.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
As explained in most macroeconomics models, capital is an important source of production and economic growth. Generating capital and making investment allows capital to flow back into the economy and fuel future output growth. Contrary to investment, saving is also an outflow of accumulated capital, which hardly has any support to economic growth in the next period. Therefore, high saving rate leads to a leftward shift in the demand curve of bond and other investment product, which ultimately results in a lower interest rate.

The theory of excess saving as an explanation to the current low interest rate in the US is not supported by domestic data, however. As #tobeupdated of figure 1# describes, though saving rate usually increases during recession where interest rate falls, the cyclical, short-run relation does not hold in longer period of time. Excluding the impact of COVID-19, both rates present long-run decreasing trend from 1982 to 2019, as interest rate fall from 13.22% to 1.55% and saving rate fall from 12.6% to 6.4%.

Global data provide a more comprehensive situation while the domestic ones fail to explain the theory. According to the World Bank data shown in #tobeupdated of figure 2#, gross saving rates in most developed countries have similar pattern as they either maintain or decrease slightly from 1982 to 2022. The world as a whole, however, experiences a slight increase in saving rate. The increase, therefore, is brought by the more significant increase of gross saving rate in developing countries, such as China, Indonesia, and Saudi Arabia. Such differentiated situation in saving rate leads to a further analysis of the relationship of developed and developing economies and their positions in the global finance market.

As the largest economy in the world, the US has engaged in the globalization of economy to a great extent, including massive trading and investing with the rest of the world. Imported $3.2 trillion and exported $2.1 trillion of goods and services in 2022, the US is the largest importer and 2nd largest exporter in the world. Meanwhile, the US is also the largest supplier and recipient of Foreign Direct Investment (FDI) among the globes, supplying 25.6% and receiving 27.8% of the world’s total amount.

Though countries in European area have always been large and important patterners of the US over centuries, most of the trade and investment end up in industrial countries. These countries share similar saving rate pattern with the US. On contrary, the connection with East Asia and Pacific (EAP) has grown rapidly during the past few decades along with accelerated globalization process. Asia-Pacific Economic Cooperation (APEC) supplies 68.8% of all imports as the member China supplies 16.8% by itself. Unlike Mexico and other developing countries in Latin America, developing countries in EAP, such as China, Korea, Indonesia, Viet Nam, and Thailand, enjoyed larger GDP growth since last century when they increased their exports to the US. Led by the 2nd largest economy, China, EAP countries are more influential to the global economy than Latin American countries. Therefore, EAP is the most suitable area to represent the global trend of developing world in the study of US economy.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References