### 2024-04-18 paper/paper.qmd
outline for paper

---
title: "An analysis of financial globalization: how excess saving in East Asian affect US interest rate"
author: Yingzhi Zhang
thanks: "Code and data are available at: https://github.com/melissazhangyz/east_asia_finance."
date: today
date-format: long
abstract: "TO BE UPDATED"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(readxl)
library(modelsummary)
```

# Introduction

# Data
## Data sourcing

## US Economy in the long run

## Economic development in East Asia and Pacific

# Model

# Result

# Discussion

\newpage

# Reference

# Discussion